---
title: Update - Integrating telemetry in SCR to improve estimation of the cost function
author: |
  | **Gates Dupont**
  | University of Massachusetts - Amherst
  | \texttt{gdupont@umass.edu}
  | 10/04/2020


header-includes:
   - \setbeamercolor{background canvas}{bg=white}
   - \setbeamertemplate{itemize item}{\small$\bullet$}
   - \setbeamertemplate{itemize subitem}{\tiny$\Box$}
   - \setbeamertemplate{itemize subsubitem}{\tiny$\blacktriangleright$}
   - \setbeamercolor{itemize item}{fg=mDarkTeal}   
   - \setbeamercolor{itemize subitem}{fg=mLightBrown}   
   - \setbeamercolor{itemize subsubitem}{fg=mDarkTeal}   
   - \usepackage{color}
   - \usepackage{verbatim}
   - \usepackage{tikz}
   - \usetikzlibrary{matrix,backgrounds,positioning}
   - \usetikzlibrary{calc}
   - \usepackage{tabto}
   - \usepackage{amsmath}
   
fontsize: 10pt
output:
 beamer_presentation:
    highlight: "default"
    theme: "metropolis"
    incremental: false
classoption: t
urlcolor: teal
keep_tex: false
---

```{r global_options, include = FALSE}
rm(list=ls())
library(knitr)
library(oSCR)
library(dplyr)
library(ggplot2)
library(raster)
library(RColorBrewer)
library(patchwork)
library(latex2exp)
library(viridis)
select = dplyr::select

opts_chunk$set(eval = FALSE, prompt = FALSE, comment = NA, 
               warning=FALSE, message=FALSE, 
               mysize=FALSE, size="\\scriptsize")

hook1 <- function(x){ gsub("```\n*```r*\n*", "", x) }
hook2 <- function(x){ gsub("```\n+```\n", "", x) }
knit_hooks$set(document = hook2)
knitr::knit_hooks$set(mysize = function(before, options, envir){
 if(before)
   return(options$size)
})
```

```{r load_data, echo = FALSE, eval=TRUE, include = FALSE}

#----Setup----

# Density
density <- 300/6400

# Tests
true_alpha2 <- rep(c(0.1, 1:5), each = 50)


#----Load sim data----

# Load SCR(-T)
attach('Valpha2_scrNOtelem_10032020.RData')
out_scr <- out %>%
  as.data.frame() %>%
  mutate(true = true_alpha2) %>%
  mutate(model = "scr-telem")
detach('file:Valpha2_scrNOtelem_10032020.RData')

# Load SCR(+T)
attach('Valpha2_test1_10012020.RData')
out_scrT <- out %>%
  as.data.frame() %>%
  mutate(true = true_alpha2) %>%
  mutate(model = "scr+telem")
detach('file:Valpha2_test1_10012020.RData')


#----Combine----

# Select common and combine
out <- rbind(
  out_scr %>%
    select(alpha2, d0, true, model),
  out_scrT %>%
    select(alpha2, d0, true, model)
)


#----Relative bias----

bias_df <- out %>% 
  group_by(true, model) %>%
  mutate(
    b_cost = 100*(alpha2 - true),
    b_density = 100*(d0 - density),
    rb_cost = 100*((alpha2 - true)/true),
    rb_density = 100*((d0 - density)/density))
```

# Background

## Model structure

__Integrated likelihood framework__

- Single model, composed of SCR and movement submodels
- Submodels share 2 params:
  - $\alpha_2$ - Cost parameter
  - $\sigma$ - Distance from home range center

- Leverages SCR data and telemetry data

## Submodels

__The SCR model__

- Basic SCR model (w/ Gaussian encounter probability model)
- Modified to estimate cost (Royle \textit{et al.} 2013)

__The movement model__ (Sutherland \textit{et al.} 2018 TWS presentation)

- Basic model of Markovian movement
  - Step length determined by LCP (ecological) distance 
  - Includes distance from activity center for SCR-like data
  - So, four params:
  

\footnotesize
|   Parameter     |          Description         |
|:---------------:|:----------------------------:|
| $\alpha_2$      | Cost value                   |
| $\sigma$        | Space use - home range       |
| $\upsilon$      | Space use - step length      |
| $\psi$          | pr(Moved)                    |
\normalsize


# $SCR_{cost}$  (base, no telem)

## Scenario settings


SCR DATA

\footnotesize
| Parameter |   Value   |           Description          |
|-----------|:---------:|:------------------------------:|
| abundance |    300    | Abundance                      |
| alpha2    | 1,2,3,4,5 | Cost value                     |
| sigma     |    0.5    | Space use - home range         |
| p0        |    0.2    | Baseline encounter probability |
| K         |     5     | Number of sampling occasions   |
\normalsize

STATE-SPACE

\footnotesize
| Parameter     |          Value          |          Description          |
|---------------|:-----------------------:|:-----------------------------:|
| rr            |         sigma/4         | State-space & cost resolution |
| lncol = lnrow |          10/rr          | State-space dimensions        |
| d0            | abundance/(lncol*lnrow) | Per-pixel density             |
\normalsize 

50 simulations per cost value



## Cost estimation

```{r scr_plot_cost, eval=TRUE, echo=FALSE}
pdf.options(encoding='CP1253.enc')

# Color palettes
blue = brewer.pal(n = 9, name = "Blues")
red = brewer.pal(n = 9, name = "Reds")
pal1 = c(red[5:9])
pal2 = c(red[4:8])

# % Bias cost
bias_df %>%
  filter(true != 0.1) %>%
  filter(model == "scr-telem") %>%
  ggplot(data = ., aes(x = as.factor(true), 
                       y = b_cost,
                       color = as.factor(true),
                       fill = as.factor(true),
                       group = as.factor(true))) +
  geom_rect(ymin = -5, ymax = 5, xmin = 0.1, xmax = 6, 
            fill = "gray70", color = "gray70") +
  geom_hline(yintercept = 0, size = 0.7) +
  geom_boxplot(lwd = 1.2) +
  scale_fill_manual(TeX("$\\alpha_2$"), values = pal2) +
  scale_color_manual(TeX("$\\alpha_2$"), values = pal1) +
  theme_minimal() + ylim(-320, 320) +
  labs(x = TeX("$\\alpha_2$"), 
       y = TeX("$\\hat{\\alpha_2}$")) +
  theme(aspect.ratio = 1, legend.position = "none",
        title = element_text(size = 25),
        text = element_text(size = 20))

```

## Density estimation

```{r scr_plot_density, eval=TRUE, echo=FALSE}

# Color palettes
blue = brewer.pal(n = 9, name = "Blues")
red = brewer.pal(n = 9, name = "Reds")
pal1 = c(red[5:9])
pal2 = c(red[4:8])

# % Bias cost
bias_df %>%
  filter(true != 0.1) %>%
  filter(model == "scr-telem") %>%
  ggplot(data = ., aes(x = as.factor(true), 
                       y = b_density,
                       color = as.factor(true),
                       fill = as.factor(true),
                       group = as.factor(true))) +
  geom_rect(ymin = -5, ymax = 5, xmin = 0.1, xmax = 6, 
            fill = "gray85", color = "gray85") +
  geom_hline(yintercept = 0, size = 0.7) +
  geom_boxplot(lwd = 1.2) +
  scale_fill_manual(TeX("$\\alpha_2$"), values = pal2) +
  scale_color_manual(TeX("$\\alpha_2$"), values = pal1) +
  theme_minimal() + ylim(-10, 10) +
  labs(x = TeX("$\\alpha_2$"), 
       y = expression(paste("% Bias in ", hat("d"[0])))) +
  theme(aspect.ratio = 1, legend.position = "none",
        title = element_text(size = 25),
        text = element_text(size = 20))

```

# $SCR_{cost}$ + telem

## Scenario settings

SCR DATA:

- Same as above


STATE-SPACE:

- Same as above
  
  
MOVEMENT DATA:

\footnotesize
| Parameter   |  Value  |          Description         |
|-------------|:-------:|:----------------------------:|
| upsilon     | sigma/4 | Space use - step length      |
| psi         |   0.7   | pr(Moved)                    |
| nfix        |  90*24  | Number of pings              |
| telemetry_n | 8       | Number of individuals tagged |
\normalsize

50 simulations per cost value


## Simulating movement data

```{r move_data, eval = TRUE, echo = FALSE}

attach('Valpha2_test1_10012020.RData')
teldata <- teldata_raw_ALL[[152]] %>%
  do.call(rbind, .) %>%
  mutate(id = rep(1:8, each = 90*24))
landscape <- landscape_ALL[[152]]
detach('file:Valpha2_test1_10012020.RData')

acs <- teldata %>%
  group_by(id) %>%
  slice(1) %>%
  as.data.frame()
  
ggplot() +
  geom_tile(data=as.data.frame(landscape, xy=T), aes(x=x, y=y, fill = layer)) +
  scale_fill_gradientn("Cost", colors = c("white", "orange3")) +
  geom_path(data = teldata, aes(x=x, y=y, group=id, color = as.factor(id)), size=0.42) +
  scale_color_manual(values = rep(brewer.pal(n = 9, name = "Blues")[6:9], 2), guide=F) +
  labs(color = NULL) +
  geom_point(data=acs, aes(x=x, y=y),
             fill = "black", color="white", pch = 21) +
  coord_equal() + theme_minimal() +
  theme(text = element_text(size=20))

```


## Cost estimation

```{r scr_telem_plot_cost, eval=TRUE, echo=FALSE}

# Color palettes
blue = brewer.pal(n = 9, name = "Blues")
red = brewer.pal(n = 9, name = "Reds")
pal1 = c(blue[5:9])
pal2 = c(blue[4:8])

# % Bias cost
bias_df %>%
  filter(true != 0.1) %>%
  filter(model == "scr+telem") %>%
  ggplot(data = ., aes(x = as.factor(true), 
                       y = b_cost,
                       color = as.factor(true),
                       fill = as.factor(true),
                       group = as.factor(true))) +
  geom_rect(ymin = -5, ymax = 5, xmin = 0.1, xmax = 6, 
            fill = "gray85", color = "gray85") +
  geom_hline(yintercept = 0, size = 0.7) +
  geom_boxplot(lwd = 1.2) +
  scale_fill_manual(TeX("$\\alpha_2$"), values = pal2) +
  scale_color_manual(TeX("$\\alpha_2$"), values = pal1) +
  theme_minimal() + ylim(-25, 25) +
  labs(x = TeX("$\\alpha_2$"), 
       y = TeX("% Bias in  $\\hat{\\alpha_2}$")) +
  theme(aspect.ratio = 1, legend.position = "none",
        title = element_text(size = 25),
        text = element_text(size = 20))

```

## Density estimation

```{r scr_telem_plot_density, eval=TRUE, echo=FALSE}

# Color palettes
blue = brewer.pal(n = 9, name = "Blues")
red = brewer.pal(n = 9, name = "Reds")
pal1 = c(blue[5:9])
pal2 = c(blue[4:8])

# % Bias cost
bias_df %>%
  filter(true != 0.1) %>%
  filter(model == "scr+telem") %>%
  ggplot(data = ., aes(x = as.factor(true), 
                       y = b_density,
                       color = as.factor(true),
                       fill = as.factor(true),
                       group = as.factor(true))) +
  geom_rect(ymin = -5, ymax = 5, xmin = 0.1, xmax = 6, 
            fill = "gray85", color = "gray85") +
  geom_hline(yintercept = 0, size = 0.7) +
  geom_boxplot(lwd = 1.2) +
  scale_fill_manual(TeX("$\\alpha_2$"), values = pal2) +
  scale_color_manual(TeX("$\\alpha_2$"), values = pal1) +
  theme_minimal() + ylim(-10, 10) +
  labs(x = TeX("$\\alpha_2$"), 
       y = expression(paste("% Bias in ", hat("d"[0])))) +
  theme(aspect.ratio = 1, legend.position = "none",
        title = element_text(size = 25),
        text = element_text(size = 20))

```

# Comparisons

## Compare cost

```{r plot_compare_cost, eval=TRUE, echo=FALSE}

# Color palettes
blue = brewer.pal(n = 9, name = "Blues")
red = brewer.pal(n = 9, name = "Reds")
pal1 = c(red[5:9], blue[5:9])
pal2 = c(red[4:8], blue[4:8])

# % Bias cost
bias_df %>%
  filter(true != 0.1) %>%
  ggplot(data = ., aes(x = as.factor(true), y = b_cost,
                       color = interaction(true, model, sep = ", "),
                       fill = interaction(true, model, sep = ", "),
                       group = interaction(true, model, sep = ", "))) +
  geom_rect(ymin = -5, ymax = 5, xmin = 0, xmax = 10, 
            fill = "gray70", color = "gray70") +
  geom_hline(yintercept = 0, size = 0.7) +
  geom_boxplot(lwd = 1.2) +
  scale_fill_manual(TeX("$\\alpha_2$, model"), values = pal2) +
  scale_color_manual(TeX("$\\alpha_2$, model"), values = pal1) +
  theme_minimal() +
  labs(x = TeX("$\\alpha_2$"), 
       y = TeX("% Bias in  $\\hat{\\alpha_2}$")) +
  theme(aspect.ratio = 1, legend.position = "none",
        title = element_text(size = 25),
        text = element_text(size = 20))

```

## Compare density

```{r plot_compare_density, eval=TRUE, echo=FALSE}

# Color palettes
blue = brewer.pal(n = 9, name = "Blues")
red = brewer.pal(n = 9, name = "Reds")
pal1 = c(red[5:9], blue[5:9])
pal2 = c(red[4:8], blue[4:8])

# % Bias cost
bias_df %>%
  filter(true != 0.1) %>%
  ggplot(data = ., aes(x = as.factor(true), y = b_density,
                       color = interaction(true, model, sep = ", "),
                       fill = interaction(true, model, sep = ", "),
                       group = interaction(true, model, sep = ", "))) +
  geom_rect(ymin = -5, ymax = 5, xmin = 0, xmax = 10, 
            fill = "gray85", color = "gray85") +
  geom_hline(yintercept = 0, size = 0.7) +
  geom_boxplot(lwd = 1.2) +
  scale_fill_manual(TeX("$\\alpha_2$, model"), values = pal2) +
  scale_color_manual(TeX("$\\alpha_2$, model"), values = pal1) +
  theme_minimal() + ylim(-8, 8) +
  labs(x = TeX("$\\alpha_2$"), 
       y = expression(paste("% Bias in ", hat("d"[0])))) +
  theme(aspect.ratio = 1, legend.position = "none",
        title = element_text(size = 25),
        text = element_text(size = 20))

```



